Why robot programming is hard

### 2014-08-25 paper.tex
Added a few more refs

We have started to see robots enter the marketplace that address two out of those three challenges. For example, iRobot's Roomba\footnote{http://www.irobot.com} only vacuums floors, but it does so in a variety of environments and is dead simple to use even by non-technical folks. On the other hand, Rethink Robotics' Baxter robot\footnote{http://www.rethinkrobotics.com/products/baxter/} can be easily programmed to perform a variety of tasks in light industrial settings, but its lack of mobility make it less suitable for the home. As another example, Willow Garage's PR-2 robot has been programmed to perform a variety of household tasks, in a variety of environments, but can only be operated by highly skilled roboticists.

@@

One important use case for personal service robots will be in-home care and assistance for disabled adults. The Robots for Humanity project~\cite{rfh}, a collaboration between Willow Garage and Georgia Tech, aimed to develop technology that enables persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Through physical therapy, Henry has regained the ability to make limited head motions (enough to control a cursor with his gaze) and to click a button with one finger.  Using an interface custom designed for him, Henry has been able to instruct Willow Garage's PR-2 robots to open cabinets in his home, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

Yet our experience with Robots for Humanity underscores how difficult it is for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door could be very difficult for Henry to accomplish. While Henry has only limited interaction capabilities, he is also strongly motivated because he has no alternatives. 
Able-bodied consumers, on the other hand, would quickly lose patience and simply fetch the towel themselves. If the state of the art in robot operation is so time-consuming, what hope do we have that ordinary mortals will be able to operate robots to perform these tasks in their own homes?

@@

\section{Robot use is programming}

@@

If you replace computer with robot, we come to a definition of robot programming: finding the right series of instructions that cause a robot to automate a task or exhibit a certain behavior. Every robot makes available a certain set of functionality -- its user-visible API. These are the primitive behaviors that have been programmed into the robot and can be activated by its user. End users must select which behavior to call when, in order to make the robot do what they want. Anyone who operates a robot, including consumers who purchase one for their home, will be programming the robot according to this definition.

@@

\begin{itemize}
\item {\bf Direct contact with the robot/task}: programming is easier if the user can see and touch the robot and its desired task. Control from a distance makes the programming process harder. For example, robots that are sent into natural disaster zones as first responders may have to be operated from a distance, which increases the programming task's difficulty because the operator's senses are limited to what can be perceived through the robot's own cameras or sensors.

\item {\bf Task/environment variance}: if the task to be performed is always the same, the programming task is easier than if the goal task changes significantly or if the environment in which the robot must operate is changing. The program must be generalized to handle this variance, which makes the programming process more difficult.

\item {\bf Goal specification}: some tasks are easy to describe, such as navigation goals specified by a single point in 3-dimensional space. Other goals, such as washing the dishes or tidying up, are more difficult to specify. Successful completion may be more dependent on how exactly the robot performs the task. Generally speaking, the more difficult it is to frame the goal specification, the more difficult the programming task will be.

\item {\bf Amount of autonomy}: robots are gaining the ability to perform some functions autonomously, such as spatial navigation and object manipulation.  The less autonomy exhibited by the robot, the easier it is to program the robot, because increased autonomy requires the programmer to develop more complex mental models of the robot's perceptions and behavior. Teleoperation interfaces are fairly easy to program because there is generally a one-to-one mapping between the user's inputs and the robot's behavior. Once a robot starts to act autonomously, understanding its behavior (and debugging it!) becomes more challenging.
\end{itemize}

>>>>

We have started to see robots enter the marketplace that address two out of those three challenges. For example, iRobot's Roomba\footnote{http://www.irobot.com} only vacuums floors, but it does so in a variety of environments and is simple to use even by non-technical folks. On the other hand, Rethink Robotics' Baxter robot\footnote{http://www.rethinkrobotics.com/products/baxter/} can be easily programmed to perform a variety of tasks in light industrial settings, but its lack of mobility makes it less suitable for the home. As another example, Willow Garage's PR-2 robot has been programmed to perform a variety of household tasks, in a variety of environments, but can only be operated by highly skilled roboticists.

Yet I believe the time is ripe for a breakthrough in personal service robotics, starting with robots that perform simpler tasks in relatively structured environments such as hospitals or hotels, and driven by better design and usable interfaces that enable them to be used by non-expert users.

@@

One important use case for personal service robots will be in-home care and assistance for disabled adults. The Robots for Humanity project~\cite{rfh}, a collaboration between Willow Garage and Georgia Tech, aimed to develop technology that enabled persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Through physical therapy, Henry had regained the ability to make limited head motions (enough to control a cursor with his gaze) and to click a button with one finger.  Using an interface custom designed for him, Henry was able to instruct Willow Garage's PR-2 robots to open cabinets in his home, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

Yet our experience with Robots for Humanity underscored how difficult it was for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door was very difficult for Henry to accomplish. While Henry had only limited interaction capabilities, he was also strongly motivated because he has no alternatives. Able-bodied consumers, on the other hand, would quickly lose patience and simply fetch the towel themselves. If the state of the art in robot operation is so time-consuming, what hope do we have that ordinary mortals will be able to operate robots to perform these tasks in their own homes?

@@

\section{Robot programming}

@@

If you replace computer with robot, we come to a definition of robot programming: finding the right series of instructions that cause a robot to automate a task or exhibit a certain behavior. Every robot makes available a certain set of functionality -- its user-visible API. These are the primitive behaviors that have been programmed into the robot and can be activated by its user. End users must select which behavior to call when, in order to make the robot do what they want.
%Anyone who operates a robot, including consumers who purchase one for their home, will be programming the robot according to this definition.

@@

\begin{itemize}
\item {\bf Direct contact with the robot/task}: programming is easier if the user can see and touch the robot and its desired task. Control from a distance makes the programming process harder~\cite{chen-ieeetos07}. For example, robots that are sent into natural disaster zones as first responders may have to be operated from a distance, which increases the programming task's difficulty because the operator's senses are limited to what can be perceived through the robot's own cameras or sensors.

\item {\bf Task/environment variance}: if the task to be performed is always the same, the programming task is easier than if the goal task changes significantly or if the environment in which the robot must operate is changing. The program must be generalized to handle this variance, which makes the programming process more difficult.
% REF: something from the EUP literature? Andy Ko or Margaret Burnett?

\item {\bf Goal specification}: some tasks are easy to describe, such as navigation goals specified by a single point in 3-dimensional space. Other goals, such as washing the dishes or tidying up, are more difficult to specify. Successful completion may be more dependent on how exactly the robot performs the task. Generally speaking, the more difficult it is to frame the goal specification, the more difficult the programming task will be.
% REF: something from the EUP literature?

\item {\bf Amount of autonomy}: robots are gaining the ability to perform some functions autonomously, such as spatial navigation and object manipulation.  The less autonomy exhibited by the robot, the easier it is to program the robot, because increased autonomy requires the programmer to develop more complex mental models of the robot's perceptions and behavior. Teleoperation interfaces are fairly easy to use because there is generally a one-to-one mapping between the user's inputs and the robot's behavior. Once a robot starts to act autonomously, understanding its behavior (and debugging it!) becomes more challenging.
\end{itemize}
